A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-07-17 tex/paper.tex
Redone Exploring posterior

%\documentclass[preprint]{aastex} 
\documentclass[iop,floatfix]{emulateapj}

@@

%\vt stands for ``vector theta''
\newcommand{\vt}{ {\bm \theta}}
\newcommand{\vT}{ {\bm \Theta}}
\newcommand{\vg}{\vt_{\star, {\rm grid}}}
\newcommand{\vpp}{\vt_{\star, {\rm post}}}
\newcommand{\finst}{f_{\lambda, {\rm inst}}}
\newcommand{\fsynth}{f_{\lambda, {\rm synth}}}
\newcommand{\vN}{\vt_{\rm N}}
\newcommand{\vtstar}{\vt_{\star}}
\newcommand{\vtcheb}{\vt_{\rm Cheb}}
\newcommand{\vtglobal}{\vt_{\rm global}}
\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
\newcommand{\vtorders}{\vt_{\rm orders}}
\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
\newcommand{\vtlines}{\vt_{\rm lines}}
\newcommand{\vtcov}{\vt_{\rm cov}}
\newcommand{\vM}{\mathsf{M}}
\newcommand{\vD}{\mathsf{D}}
\newcommand{\vR}{\mathsf{R}}
\newcommand{\vC}{\mathsf{C}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

@@

@@

There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
However, in general this approach is still computationally prohibitive for any iterative method of 
probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
structures that were pre-computed for a discrete grid of parameter values, $\{\vt_{\ast}\}^{\rm 
grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
that interpolated atmosphere to synthesize $f_{\lambda}$ \citep[e.g., as for SME;][]{valenti96}.  A more 
common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
technically the former approach is most similar to the ideal case, the computational cost of 
repeated spectral synthesis is sufficiently high to make a detailed exploration of parameter space 
(particularly for data with a large spectral range) considerably less appealing.  A related, but 
different approach is to eschew forward modeling entirely (and therefore repeated spectral 
syntheses and/or library interpolations), and instead evaluate the models only at the discrete grid 
points of the library.  Then, these discretized samples of the posterior probability density can be 
interpolated to an arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals 
\citep[e.g., the method of SPC;][]{buchhave12}.  The difficulty with this latter approach is that the 
parameter uncertainties can be smaller than the grid spacing; in that case, there is valid concern 
that this interpolation might not accurately recover intrinsic parameter degeneracies.

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$f_{\lambda}(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by the iron abundance, [Fe/H]).  However, 
it is worth noting that the techniques we will develop are applicable to {\it any} approach for 
generating a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
f_{\lambda}(\{\vt_{\ast}\}^{\rm grid}) \leadsto f_{\lambda}(\vt_{\ast}),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).

@@

Generally, the ``raw" model spectrum $f_{\lambda}(\vt_{\ast})$ will be highly over-sampled compared 
to a typical observed spectrum, and does not account for several additional observational and 
instrumental effects that become important in comparisons with real data.  Therefore, a certain 
amount of post-processing is required before assessing the model quality.  We treat that 
post-processing in two stages: the first deals with an additional set of ``observational" 
parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
location of the target, while the second employs a suite of nuisance (hyper-)parameters, $\vT_{\rm 
n}$, designed to mitigate an imperfect data calibration.

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,
\begin{equation}
f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) = f_{\lambda}(\vt_{\ast}) \otimes \varphi_v = f_{\lambda}(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
f_{\lambda}(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-dimensional array $\vM$ (and $N_{\rm pix}$ is the number of pixels in the spectrum).  
For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.

@@

We adopt a somewhat analogous approach to deal with this issue, but cast it internal to the 
modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
the model spectrum, enabling us to explore the distribution of possible calibrations (see 
Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
In essence, the inferences for the relevant stellar parameters will properly account for the 
underlying uncertainty in the calibration process.  This is achieved in practice by distorting the 
model spectrum with a (low-order) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
\begin{equation}
\vM(\vt, \vT_{\rm cheb}) = \vM(\vt) \times \sum_n c_n \, T_n,
\label{eqn:chebyshev}
\end{equation}
where $T_n$ are the standard Chebyshev functions of order $n$ and the coefficients are treated as a
set of nuisance (hyper-)parameters, $\vT_{\rm cheb} = \{c_n\}^{\rm order}$, for each available 
spectral order in the dataset.  With judicious priors on $\vT_{\rm cheb}$, we can ensure that the 
unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-order (scaling) 
coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
Therefore, we enforce an additional constraint by requiring that the mean of the polynomial is 
unity.  For data with a single spectral order, this means setting $c_0 = 1$.  If the goal is to 
model multiple spectral orders, we assign $c_0 = 1$ in an arbitrary order as an anchor, but permit 
the $c_0$ values in other orders to be different (as necessary).  It is worth noting that this 
formalism can, in principle, be extended to develop models for completely uncalibrated spectra, 
rather than the residual calibration mismatches as described here (with a suitable relaxation of 
the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how 
these nuisance parameters are applied.

@@

The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding 
$N_{\rm pix}$-dimensional residual spectrum can be defined for any input parameter set,
\begin{equation}
\vR \equiv \vR(\vt, \vT_{\rm cheb}) \equiv \vD-\vM(\vt, \vT_{\rm cheb}).
\end{equation}
To quantify the probability of the data conditioned on the model, we adopt a standard 
multi-dimensional Gaussian likelihood function,
\begin{equation}
p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2}
   \vR^\trans \vC^{-1} \vR \right ), 
   \label{eqn:prob}
\end{equation}
which penalizes models that yield larger residuals and explicitly allows for covariances in the 
residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix, $\vC$.  For practical 
numerical reasons, we generally consider the log-likelihood as the fit quality metric, which in 
this case is
\begin{equation}
  \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
  \label{eqn:lnprob}
\end{equation}

The covariance matrix $\vC$ characterizes both the measurement uncertainty ($\sigma$; 
``noise") in each pixel and the intrinsic covariance between pixels.  The special case where each 
pixel represents an independent measurement results in a diagonal covariance matrix, $\vC_{ij} = 
\delta_{ij} \,\, \sigma_i$ where $\sigma_i$ is the uncertainty in pixel $i$ and $\delta_{ij}$ is 
the Kronecker delta function, and Eq.~\ref{eqn:lnprob} reduces to the familiar
\begin{equation}
\ln{p(\vD | \vM)} = -\frac{1}{2} \sum_i^{N_{\rm pix}} \frac{\vR_i^2}{\sigma_i^2} \equiv -\frac{\chi^2}{2},
\label{eqn:chisq}
\end{equation}
the sum of the square of the residuals weighted by the inverse variances (squared uncertainties).  
However, the problem being addressed here necessitates the use of a more complex covariance matrix; 
additional off-diagonal terms that can explicitly characterize (1) pixel-to-pixel covariances 
imposed by the discrete over-sampling of the line-spread function, and (2) highly correlated 
residuals as manifestations of the still-imperfect model library are required to avoid biasing our 
inferences of the physically interesting parameters ($\vt$).  The following subsections describe 
how these issues are addressed in the practical implementation of $\vC$.

@@

For a well-designed spectrograph and sufficiently accurate model spectrum, this
{\it global} (i.e., present throughout the spectrum) covariant structure should
have a relatively low amplitude and small correlation length.  To describe that
structure, we assume a stationary covariance kernel (also called a radial basis
function) with an amplitude that depends only on the velocity distance between
two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  The kernel is chosen to parametrically describe the covariance 
between pixel residuals, such that
\begin{equation}
  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian
processes to parameterize such a mildly covariant structure
\citep[e.g.,][]{rasmussen05}.  Here we adopt the \matern\ kernel with 
$\nu = 3/2$,
\begin{equation}
  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
which is parameterized by an amplitude $a_g$ and scale $\ell$ and makes for a smooth transition to 
negligible covariance at large $r$.  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expendiency, we employ a Hann window function
\begin{equation}
  w\,(\lambda_i, \lambda_j |\, r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
to taper the kernel (so the effective kernel is $w \, \mathcal{K}_g$).  The truncation distance 
$r_0$ can be fixed to some reasonable multiple of the scale parameter (here we set $r_0 = 4\ell$).  
The examples in Figure \ref{fig:matern} demonstrate that such a kernel readily produces correlated 
structure in the residual spectrum that is similar to those seen in a typical data--model 
comparison.

@@

We can now compute the covariance matrix employed in the likelihood calculation 
(Eq.~\ref{eqn:lnprob}) as the linear combination of these kernels and the trivial pixel-by-pixel 
noise matrix, 
\begin{multline}
\vC_{ij}(\vT_{\rm cov})  = b \, \delta_{ij}\sigma_i + \\
         w(r_{ij}|r_0 = 4\ell) \,\, \mathcal{K}_g(\lambda_i, \lambda_j|a_g, \ell) + \\ 
         \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \mathcal{K}_k(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
\end{multline}
where the covariance hyperparameters $\vT_{\rm cov} = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
h_k\}^{N_{\rm reg}}]$, $N_{\rm reg}$ is the number of local residual regions
(see below for details on how this is determined), and $b$ is assumed to be a
fixed parameter that scales up the pixel noise values to account for read
noise, noise added during the 2D to 1D spectral extraction procedure, and
interpolation uncertainties (see Section \ref{subsec:synthetic}; reasonable
values are $b \approx 1.02$--1.10 for well-calibrated optical spectra; see Section
\ref{sec:examples} for examples).

>>>>

\documentclass[preprint]{aastex} 
%\documentclass[iop,floatfix]{emulateapj}

@@

\newcommand{\vM}{\mathsf{M}}
\newcommand{\vD}{\mathsf{D}}
\newcommand{\vR}{\mathsf{R}}
\newcommand{\vC}{\mathsf{C}}
\newcommand{\fM}{ \vec{{\bm M}}}
\newcommand{\fMi}{M_i}
\newcommand{\fD}{ \vec{{\bm D}}}
\newcommand{\fDi}{D_i}
\newcommand{\fR}{ {\bm R}}
\newcommand{\dd}{\,{\rm d}}
\newcommand{\trans}{\mathsf{T}}
\newcommand{\Z}{[{\rm Fe}/{\rm H}]}
\newcommand{\A}{[\alpha/{\rm Fe}]}
\newcommand{\matern}{Mat\'{e}rn}

%flux
\newcommand{\flam}{f_\lambda}

%\vt stands for ``vector theta''
\newcommand{\vt}{ {\bm \theta}}
\newcommand{\vT}{ {\bm \Theta}}

%\vp stands for ``vector phi''
\newcommand{\vp}{ {\bm \phi}}
\newcommand{\vP}{ {\bm \Phi}}

\newcommand{\fundamentalParameters}{\vt_{\ast}}
\newcommand{\gridParameters}{\{\fundamentalParameters\}^{\rm 
grid}}
\newcommand{\observationalParameters}{\vt_{\textrm{obs}}}
\newcommand{\stellarParameters}{\vt}

%\phi and \Phi are used for nuisance parameters
\newcommand{\cheb}{ \vp_{\textrm{Cheb}}}
\newcommand{\chebi}[1]{ \vp_{\textrm{Cheb}_{#1}}} %for signifying a specific order
\newcommand{\Cheb}{ \vP_{\textrm{Cheb}}}
\newcommand{\Chebi}[1]{ \vP_{\textrm{Cheb}_{\ne #1}}} %for specifying everything but a specific order
\newcommand{\cov}{ \vp_{\textrm{cov}}}
\newcommand{\covi}[1]{ \vp_{\textrm{cov}_{#1}}} %for signifying a specific order
\newcommand{\Cov}{ \vP_{\textrm{cov}}}
\newcommand{\Covi}[1]{ \vP_{\textrm{cov}_{\ne #1}}} %for specifying everything but a specific order

\newcommand{\allParameters}{\vT} %all parameters
\newcommand{\nuisanceParameters}{\vP} %all nuisance parameters

%kernel notation
\newcommand{\KK}{\mathcal{K}}
\newcommand{\Kglobal}{\KK_{\textrm{g}}}
\newcommand{\Klocal}{\KK_k}

@@

\newcommand{\vtcheb}{\vt_{\rm Cheb}}
\newcommand{\vtglobal}{\vt_{\rm global}}
\newcommand{\vtorder}[1]{\vt_{{\rm order}_{#1}}}
\newcommand{\vtorders}{\vt_{\rm orders}}
\newcommand{\vtline}[1]{\vt_{{\rm line}_{#1}}}
\newcommand{\vtlines}{\vt_{\rm lines}}
\newcommand{\vtcov}{\vt_{\rm cov}}

@@

There are various approaches to synthesizing a spectrum, $\flam$, for a specific set of model
parameters, $\fundamentalParameters$.  In an ideal case, a model stellar atmosphere is constructed and then 
subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
However, in general this approach is still computationally prohibitive for any iterative method of 
probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
structures that were pre-computed for a discrete grid of parameter values, $\gridParameters$, for some arbitrary $\fundamentalParameters$, and then perform a radiative transfer calculation with 
that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for SME;][]{valenti96}.  A more 
common variant is to instead rely on interpolation over a library of pre-synthesized model spectra, 
$\flam(\gridParameters)$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While 
technically the former approach is most similar to the ideal case, the computational cost of 
repeated spectral synthesis is sufficiently high to make a detailed exploration of parameter space 
(particularly for data with a large spectral range) considerably less appealing.  A related, but 
different approach is to eschew forward modeling entirely (and therefore repeated spectral 
syntheses and/or library interpolations), and instead evaluate the models only at the discrete grid 
points of the library.  Then, these discretized samples of the posterior probability density can be 
interpolated to an arbitrary $\fundamentalParameters$ to construct appropriate confidence intervals 
\citep[e.g., the method of SPC;][]{buchhave12}.  The difficulty with this latter approach is that the 
parameter uncertainties can be smaller than the grid spacing; in that case, there is valid concern 
that this interpolation might not accurately recover intrinsic parameter degeneracies.

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$\flam(\gridParameters)$, where $\fundamentalParameters = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by the iron abundance, [Fe/H]).  However, 
it is worth noting that the techniques we will develop are applicable to {\it any} approach for 
generating a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
\flam(\gridParameters) \leadsto \flam(\fundamentalParameters),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\gridParameters$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\gridParameters$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).

@@

Generally, the ``raw" model spectrum $\flam(\fundamentalParameters)$ will be highly over-sampled compared 
to a typical observed spectrum, and does not account for several additional observational and 
instrumental effects that become important in comparisons with real data.  Therefore, a certain 
amount of post-processing is required before assessing the model quality.  We treat that 
post-processing in two stages: the first deals with an additional set of ``observational" 
parameters, $\observationalParameters$, that incorporate dynamical effects, geometry, and the relative 
location of the target, while the second employs a suite of nuisance (hyper-)parameters, $\Cheb$, designed to mitigate an imperfect data calibration.

We can further divide $\observationalParameters$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\observationalParameters = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, treats the instrumental 
spectral broadening.  For illustrative purposes we assume $\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,
\begin{equation}
\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) = \flam(\fundamentalParameters) \otimes \varphi_v = \flam(\fundamentalParameters) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
\flam(\fundamentalParameters, \vt_{{\rm obs}, v}) \mapsto \vM(\fundamentalParameters, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-dimensional array $\vM$ (and $N_{\rm pix}$ is the number of pixels in the spectrum).  
For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.

@@

We adopt a somewhat analogous approach to deal with this issue, but cast it internal to the 
modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
the model spectrum, enabling us to explore the distribution of possible calibrations (see 
Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
In essence, the inferences for the relevant stellar parameters will properly account for the 
underlying uncertainty in the calibration process.  For a continuous spectrum,
such as a long-slit spectrum or a single spectral order of an echelle spectrum,
this is achieved in practice by distorting the model spectrum with a
(low-degree) Chebyshev polynomial \citep[e.g.,][]{eisenstein06,koleva09}, 
\begin{equation}
\vM(\stellarParameters, \cheb) = \vM(\stellarParameters) \times \sum_n c_n \, T_n,
\label{eqn:chebyshev}
\end{equation}
where $T_n$ are the standard Chebyshev functions of degree $n$ and the coefficients are treated as a
set of nuisance (hyper-)parameters, $\cheb = \{c_0, c_1, \ldots, c_{n-1} \}$. For an echelle spectrum with $N$ spectral orders, there is a Chebsyhev polynomial for each order of the spectrum $i$, such that the total collection of Chebyshev parameters is $\Cheb = \{\chebi{1}, \chebi{2}, \ldots, \chebi{N} \}$.  With judicious priors on $\Cheb$, we can ensure that the 
unintended treatment of real spectral features (e.g., broad, deep molecular bands) as calibration 
artifacts is negligible (see Section~\ref{sec:examples} for examples).  The lowest-degree (scaling) 
coefficient, $c_0$, is by its nature degenerate with the solid angle parameter, $\Omega$.  
Therefore, we enforce an additional constraint by requiring that the mean of the polynomial is 
unity.  For data with a single spectral order, this means setting $c_0 = 1$.  If the goal is to 
model multiple spectral orders, we assign $c_0 = 1$ in an arbitrary order as an anchor, but permit 
the $c_0$ values in other orders to be different (as necessary).  It is worth noting that this 
formalism can, in principle, be extended to develop models for completely uncalibrated spectra, 
rather than the residual calibration mismatches as described here (with a suitable relaxation of 
the priors on $\vT_{\rm cheb}$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how 
these nuisance parameters are applied.

@@

The quality of the model spectrum is assessed by comparing to the data with a pixel-by-pixel 
likelihood calculation.  If we denote the data spectrum as $\vD$, then a corresponding 
$N_{\rm pix}$-dimensional residual spectrum can be defined for any input parameter set,
\begin{equation}
\vR \equiv \vR(\stellarParameters, \cheb) \equiv \vD-\vM(\stellarParameters, \cheb).
\end{equation}
To quantify the probability of the data conditioned on the model, we adopt a standard 
multi-dimensional Gaussian likelihood function,
\begin{equation}
p(\vD|\vM) =  \frac{1}{[(2 \pi)^{N_{\rm pix}} \det(\vC)]^{1/2}} \exp\left ( -\frac{1}{2}
   \vR^\trans \vC^{-1} \vR \right ), 
   \label{eqn:likelihood}
\end{equation}
which penalizes models that yield larger residuals and explicitly allows for covariances in the 
residual spectrum through the $N_{\rm pix} \times N_{\rm pix}$ matrix, $\vC$.  For practical 
numerical reasons, we generally consider the log-likelihood as the fit quality metric, which in 
this case is
\begin{equation}
  \ln{p(\vD | \vM)} = -\frac{1}{2} \left( \vR^\trans \vC^{-1} \vR + \ln{\det{\vC}} + N_{\rm pix} \ln{2\pi} \right).
  \label{eqn:lnlikelihood}
\end{equation}

The covariance matrix $\vC$ characterizes both the measurement uncertainty ($\sigma$; 
``noise") in each pixel and the intrinsic covariance between pixels.  The special case where each 
pixel represents an independent measurement results in a diagonal covariance matrix, $\vC_{ij} = 
\delta_{ij} \,\, \sigma_i$ where $\sigma_i$ is the uncertainty in pixel $i$ and $\delta_{ij}$ is 
the Kronecker delta function, and Eq.~\ref{eqn:lnlikelihood} reduces to the familiar
\begin{equation}
\ln{p(\vD | \vM)} = -\frac{1}{2} \sum_i^{N_{\rm pix}} \frac{\vR_i^2}{\sigma_i^2} \equiv -\frac{\chi^2}{2},
\label{eqn:chisq}
\end{equation}
the sum of the square of the residuals weighted by the inverse variances (squared uncertainties).  
However, the problem being addressed here necessitates the use of a more complex covariance matrix; 
additional off-diagonal terms that can explicitly characterize (1) pixel-to-pixel covariances 
imposed by the discrete over-sampling of the line-spread function, and (2) highly correlated 
residuals as manifestations of the still-imperfect model library are required to avoid biasing our 
inferences of the physically interesting parameters ($\vt$).  The following subsections describe 
how these issues are addressed in the practical implementation of $\vC$.

@@

For a well-designed spectrograph and sufficiently accurate model spectrum, this
{\it global} (i.e., present throughout the spectrum) covariant structure should
have a relatively low amplitude and small correlation length.  To describe that
structure, we assume a stationary covariance kernel (also called a radial basis
function) with an amplitude that depends only on the velocity distance between
two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  The kernel is chosen to parametrically describe the covariance 
between pixel residuals, such that
\begin{equation}
  \Kglobal(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian
processes to parameterize such a mildly covariant structure
\citep[e.g.,][]{rasmussen05}.  Here we adopt the \matern\ kernel with 
$\nu = 3/2$,
\begin{equation}
  \Kglobal(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
which is parameterized by an amplitude $a_g$ and scale $\ell$ and makes for a smooth transition to 
negligible covariance at large $r$.  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expendiency, we employ a Hann window function
\begin{equation}
  w\,(\lambda_i, \lambda_j |\, r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
to taper the kernel (so the effective kernel is $w \, \Kglobal$).  The truncation distance 
$r_0$ can be fixed to some reasonable multiple of the scale parameter (here we set $r_0 = 4\ell$).  
The examples in Figure \ref{fig:matern} demonstrate that such a kernel readily produces correlated 
structure in the residual spectrum that is similar to those seen in a typical data--model 
comparison.

@@

We can now compute the covariance matrix employed in the likelihood calculation 
(Eq.~\ref{eqn:lnlikelihood}) as the linear combination of the trivial pixel-by-pixel 
noise matrix and these kernels, 
\begin{multline}
\vC_{ij}(\cov)  = b \, \delta_{ij}\sigma_i + \\
         w(r_{ij}|r_0 = 4\ell) \,\, \Kglobal(\lambda_i, \lambda_j|a_g, \ell) + \\ 
         \sum_k w(r_{ij}|r_0 = 4\sigma_k) \,\, \mathcal{S}_k(\lambda_i,\lambda_j | h_k) \,\, \Klocal(\lambda_i, \lambda_j | a_k, \mu_k, \sigma_k),
\end{multline}
where the covariance hyperparameters $\cov = [a_g, \ell, \{a_k, \mu_k, \sigma_k, 
h_k\}^{N_{\rm reg}}]$, $N_{\rm reg}$ is the number of local residual regions
(see below for details on how this is determined), and $b$ is assumed to be a
fixed parameter that scales up the pixel noise values to account for read
noise, noise added during the 2D to 1D spectral extraction procedure, and
interpolation uncertainties (see Section \ref{subsec:synthetic}; reasonable
values are $b \approx 1.02$--1.10 for well-calibrated optical spectra; see Section
\ref{sec:examples} for examples). If the echelle spectrum contains $N$ spectral orders, then the collection of covariance parameters for the whole spectrum are $\Cov = \{\covi{1}, \covi{2}, \ldots, \covi{N} \}$.
